---
title: "Analyzing Big Data with Microsoft R "
output: html_document
---
### Introduction 
2016 Taxi data from NY.
http://www.nyc.gov/html/tlc/html/about/trip_record_data.shtml


### Data Analysis; Read, Explore, Transform and Model the data

### Reading Data
#### Read a portion of data into Data Frame that will fit into memory
10% of one month (January)

```{r}
rxSetComputeContext("local")
rxGetComputeContext()

#input_csv <- './Data/yellow_tripsample_2016-01.csv'
#colClasses = col_classes
input_csv <- './Data/nyc_sample.csv'
nyc_sample_df <- read.csv(input_csv, header = T, stringsAsFactors = F)
names(nyc_sample_df)[1] <- "VendorID"
#nyc_sample_df <- read.csv(input_csv, nrows = 1000, header = TRUE, sep = ",", quote = "\"", stringsAsFactors = FALSE)
nyc_sample_df$VendorID <- as.factor(nyc_sample_df$VendorID)
nyc_sample_df$RatecodeID <- as.factor(nyc_sample_df$RatecodeID)
nyc_sample_df$store_and_fwd_flag <- as.factor(nyc_sample_df$store_and_fwd_flag)
dim(nyc_sample_df)
head(nyc_sample_df)
str(nyc_sample_df)

```

#### Read data into local compute context from a XDF (External file Format file)

This is ~70 million rows (2.9 Gig) with 20 columns representing six months worth of 2016 data (January through June). 
It took 23 minutes to build the XDF (eXternal Data File) 

```{r}

rxOptions(reportProgress = 0)
input_xdf <- './Data/nyc_full.xdf'
nyc_full_xdf <- RxXdfData(input_xdf)
rxGetInfo(nyc_full_xdf)

system.time(
  rxsumfull_xdf <- rxSummary( ~., nyc_full_xdf) # provide statistical summaries for fare amount
)
rxsumfull_xdf

```
Import a sample (3,467,953 rows) from a csv into a xdf 

```{r}

col_classes <- c('VendorID' = "factor",
                 'tpep_pickup_datetime' = "character",
                 'tpep_dropoff_datetime' = "character",
                 'passenger_count' = "integer",
                 'trip_distance' = "numeric",
                 'pickup_longitude' = "numeric",
                 'pickup_latitude' = "numeric",
                 'RatecodeID' = "factor",
                 'store_and_fwd_flag' = "factor",
                 'dropoff_longitude' = "numeric",
                 'dropoff_latitude' = "numeric",
                 'payment_type' = "factor",
                 'fare_amount' = "numeric",
                 'extra' = "numeric",
                 'mta_tax' = "numeric",
                 'tip_amount' = "numeric",
                 'tolls_amount' = "numeric",
                 'improvement_surcharge' = "numeric",
                 'total_amount' = "numeric",
                 'u' = "numeric")


nyc_sample_csv <- './Data/nyc_sample.csv'
nyc_sample_xdf<-'./Data/nyc_sample.xdf'
input_xdf <- './Data/nyc_sample'
input_xdf <- './Data/yellow_tripdata_2016.xdf'
rxImport(nyc_sample_csv, nyc_sample_xdf, rowSelection = (payment_type != 0), overwrite = T, colClasses = col_classes)
system.time(
  rxsumsamp_xdf <- rxSummary(~payment_type, nyc_sample_xdf) # provide statistical summaries for fare amount
)
rxsumsamp_xdf

#Memory Sizes
sprintf('Sample Dataframe Memory footprint = %.2f Meg', object.size(nyc_sample_df) / 1000000)
sprintf('Full XDF File Memory footprint = %.2f bytes', object.size(nyc_full_xdf))
sprintf('Sample XDF Memory footprint = %.2f bytes', object.size(nyc_sample_xdf))
```

#### Use a .csv with RevoScale R

This is an example of using RX functions to read a regular csv file into a Revoscale format in memory.

```{r}

input_csv <- './Data/yellow_tripsample_2016-01.csv' # we can only use one month's data unless we join the CSVs
nyc_csv <- RxTextData(input_csv, colClasses = col_classes) # point to CSV file and provide column info
system.time(
  rxsum_csv <- rxSummary( ~ payment_type, nyc_csv) # provide statistical summaries for fare amount
)
rxsum_csv

```
### Data Exploration
#### str and head command all in one
#### Accessing rx S4 objects

```{r,echo=TRUE}

nyc_xdfInfo <- rxGetVarInfo(nyc_sample_xdf)
str(nyc_xdfInfo)
names(nyc_xdfInfo$tip_amount)
nyc_xdfInfo$tip_amount
typeof(nyc_xdfInfo$tip_amount)
length(nyc_xdfInfo$tip_amount)
nyc_xdfInfo$tip_amount[[4]]
nyc_xdfInfo$tip_amount[["high"]]

```
#### Some Graphs

```{r, fig.width=8, fig.height=6}

system.time(
rxHistogram( ~ trip_distance, nyc_full_xdf, startVal = 0, endVal = 25, histType = "Percent", numBreaks = 20, reportProgress = 0)
    )

```

```{r, fig.width=4, fig.height=6}

system.time(
rxHistogram( ~ trip_distance, nyc_full_xdf, startVal = 1000, endVal = 10000, histType = "Percent", numBreaks = 20, reportProgress = 0
    , rowSelection = (trip_distance > 1000 & trip_distance < 10000))
    )

```

```{r, fig.width=4, fig.height=6}

system.time(
rxHistogram( ~ trip_distance, nyc_full_xdf, startVal = 1000, endVal = 10000, histType = "Counts", numBreaks = 20, reportProgress = 0
    , rowSelection = (trip_distance > 1000 & trip_distance < 10000))
    )

```

```{r, fig.width=4, fig.height=6}

system.time(
rxHistogram( ~ fare_amount, nyc_full_xdf, startVal = 1000, endVal = 10000, histType = "Counts", numBreaks = 20, reportProgress = 0
    , rowSelection = (fare_amount > 1000 & fare_amount < 10000))
    )

```

### Data Clean-up

```{r}

system.time(
rxDataStep(inData = nyc_full_xdf, outFile = './Data/nyc_clean.xdf'
    , rowSelection = (trip_distance > 0 & trip_distance < 1000)
    & (fare_amount > 0 & fare_amount < 1000)
    & (tip_amount >= 0)
    & (tolls_amount >= 0)
    & (total_amount >= 0 & total_amount < 1000)
    & (mta_tax >= 0)
    & (extra >= 0)
    , overwrite = T)
    )
nyc_clean_xdf <- RxXdfData('./Data/nyc_clean.xdf')
system.time(
clean<-rxSummary( ~ ., data = nyc_clean_xdf)
)
clean 

```

### Feature Engineering
#### Performing simple and complex transformations

```{r}
payment_labels <- c('Credit card', 'Cash')
system.time(
rxDataStep(nyc_clean_xdf,nyc_clean_xdf,
           transforms = list(Ratecode_type_desc = factor(RatecodeID, levels = 1:6, labels = c('Standard rate', 'JFK', 'Newark', 'Nassau or Westchester', 'Negotiated fare', 'Group ride'))
           , payment_type_desc = factor(payment_type, levels = 1:2, labels = payment_labels)),
           transformObjects = list(payment_labels = payment_labels),
           overwrite = TRUE)
	)
system.time(
ratecode<- rxSummary( ~ Ratecode_type_desc, nyc_clean_xdf)
)
ratecode

system.time(
payment<-rxSummary( ~ payment_type_desc, nyc_clean_xdf)
)
payment 
```

### Change Context


```{r, echo=TRUE}


rxSetComputeContext("local")
rxGetComputeContext()

sqlConnString <- "Driver=SQL Server;Server=localhost;Database=taxi;trusted_connection=true;Integrated Security=SSPI;"

sqlRowsPerRead <- 1000000

sqlTable <- 'yellow_manhattan'

ccColInfo <- list(
  tpep_pickup_datetime = list(type = "character"),
  tpep_dropoff_datetime = list(type = "character"),
  passenger_count = list(type = "integer"),
  trip_distance = list(type = "numeric"),
  pickup_longitude = list(type = "numeric"),
  pickup_latitude = list(type = "numeric"),
  dropoff_longitude = list(type = "numeric"),
  dropoff_latitude = list(type = "numeric"),
  RateCodeID = list(type = "factor", levels = as.character(1:6), newLevels = c("standard", "JFK", "Newark", "Nassau or Westchester", "negotiated", "group ride")),
  store_and_fwd_flag = list(type = "factor", levels = c("Y", "N")),
  payment_type = list(type = "factor", levels = as.character(1:2), newLevels = c("card", "cash")),
  fare_amount = list(type = "numeric"),
  tip_amount = list(type = "numeric"),
  total_amount = list(type = "numeric")
)

weekday_labels <- c('Sun', 'Mon', 'Tue', 'Wed', 'Thu', 'Fri', 'Sat')
hour_labels <- c('1AM-5AM', '5AM-9AM', '9AM-12PM', '12PM-4PM', '4PM-6PM', '6PM-10PM', '10PM-1AM')

ccColInfo$pickup_dow <- list(type = "factor", levels = weekday_labels)
ccColInfo$pickup_hour <- list(type = "factor", levels = hour_labels)
ccColInfo$dropoff_dow <- list(type = "factor", levels = weekday_labels)
ccColInfo$dropoff_hour <- list(type = "factor", levels = hour_labels)

nyc_shapefile <- readShapePoly('./Data/ZillowNeighborhoods-NY/ZillowNeighborhoods-NY.shp')
mht_shapefile <- subset(nyc_shapefile, str_detect(CITY, 'New York City-Manhattan'))
manhattan_nhoods <- as.character(mht_shapefile@data$NAME)

ccColInfo$pickup_nhood <- list(type = "factor", levels = manhattan_nhoods)
ccColInfo$dropoff_nhood <- list(type = "factor", levels = manhattan_nhoods)

#RxXDFData
#RxSqlServerData

nyc_sql <- RxSqlServerData(connectionString = sqlConnString, table = sqlTable, rowsPerRead = sqlRowsPerRead, colInfo = ccColInfo)

rxGetInfo(nyc_sql, getVarInfo = TRUE, numRows = 3) # show column types and the first 10 rows

system.time(
  rxsum_sql <- rxSummary( ~ ., nyc_sql) # provide statistical summaries for all the columns
)
rxsum_sql

rxGetComputeContext()

sqlWait <- TRUE
sqlConsoleOutput <- FALSE
sqlShareDir <- getwd()

sqlComputeTrace <- RxInSqlServer(
    connectionString = sqlConnString,
    shareDir = sqlShareDir,
    , wait = sqlWait
    , consoleOutput = sqlConsoleOutput
    #,traceEnabled = TRUE
    #,traceLevel = 7
    )

rxSetComputeContext(sqlComputeTrace)
#rxSetComputeContext("local")

rxGetComputeContext()

nyc_sql <- RxSqlServerData(connectionString = sqlConnString, table = sqlTable, rowsPerRead = sqlRowsPerRead, colInfo = ccColInfo)

rxGetInfo(nyc_sql, getVarInfo = TRUE, numRows = 3) # show column types and the first 10 rows

system.time(
  rxsum_sql <- rxSummary( ~ ., nyc_sql) # provide statistical summaries for all the columns
)
rxsum_sql

```

### Model

```{r}

rxGetComputeContext()

sql_trn <- 'SELECT * FROM [dbo].[yellow_manhattan] WHERE u < .035' #65%

nyc_sql_trn <- RxSqlServerData(connectionString = sqlConnString, sqlQuery = sql_trn
    , rowsPerRead = sqlRowsPerRead, colInfo = ccColInfo)

sql_tst <- 'SELECT * FROM [dbo].[yellow_manhattan] WHERE u >=.035' #35%
nyc_sql_tst <- RxSqlServerData(connectionString = sqlConnString, sqlQuery = sql_tst
    , rowsPerRead = sqlRowsPerRead, colInfo = ccColInfo)

rxGetInfo(nyc_sql_tst, getVarInfo = TRUE, numRows = 3) # show column types and the first 10 rows

system.time(
  rxsum_sql <- rxSummary( ~ ., nyc_sql) # provide statistical summaries for all the columns
)
rxsum_sql

system.time(linmod <- rxLinMod(tip_percent ~ pickup_nhood:dropoff_nhood + pickup_dow:pickup_hour,
                               data = nyc_sql_trn, reportProgress = 0)
)

```
### Questions