---
title: "mlr3 Basics"
#author: "Matt Dube"
date: "9/29/2019 (updated: `r Sys.Date()`)"
output: 
    html_document:
      toc: true
      toc_float: true
      toc_depth: 3
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Setup

Load basic libraries and data. 

We'll be using the Employee Attrition dataset from the rsample package.

```{r setup_part_one}
library(ggplot2)
library(ggthemes)
library(data.table)

library(mlr3)
library(mlr3viz)

attrDT <- fread(here::here("00_Data/", "attrition_train.csv"))
attrDT[,Attrition:=as.factor(Attrition)]

#attrDT <- as.data.table(rsample::attrition)
```

### review data

```{r setup_part_two}
attrDT[,1:5]

table(attrDT$Attrition)

```

## mlr3 workflow
The list of built-in learners (algorithms) is stored in the R6 dictionary mlr_learers.

```{r}
as.data.table(mlr_learners)
```

<br>
You can load additional learners using the <font face=bold color="red">mlr3learners</font> package.

```{r more_learners}
library(mlr3learners)

as.data.table(mlr_learners)[,.(key,predict_types,packages)]

```

<br>
This is a classification problem. We're going to next create a task and define a couple of learners.

### Create Task

```{r create_task}

tsk_attr <- TaskClassif$new(id="attrition", backend = attrDT, 
                            target = "Attrition", positive = "Yes")
tsk_attr
    
```

Review task properties.

```{r}
tsk_attr$data()[,1:4]

tsk_attr$feature_names
```

There is quite a bit of information available on a task. Here are some examples, but you should explore and review others.

- `tsk_attr$feature_names`
- `tsk_attr$feature_types`
- `tsk_attr$target_names`
- `tsk_attr$ncol`
- `tsk_attr$nrow`
- `tsk_attr$missings()`

### Create Learners

```{r create_learners}
lrn_dummy  <- lrn("classif.featureless")
lrn_rpart  <- lrn("classif.rpart")
lrn_ranger <- lrn("classif.ranger")

```

<br>
Review the learners:
<br>
```{r lrn_review}
lrn_dummy
lrn_rpart
lrn_ranger
```

Now that we have a task (data), and some learners (algrithms), we can split our data and start training models.

### Data Split
mlr3 splits data by creating train and test indices. Separate datasets are not created.

```{r split_data}
set.seed(4411)
train.idx <- sample(seq_len(tsk_attr$nrow), 0.7 * tsk_attr$nrow)
test.idx <- setdiff(seq_len(tsk_attr$nrow), train.idx)

```

### Train Models
And now we can train a model for each learner using this basic syntax:

`learner$train(task, data)`
<br>
```{r train_dummy}
lrn_dummy$train(tsk_attr, row_ids = train.idx)

```

Let's compare the rpart learner before and after training:

```{r train_rpart}
lrn_rpart
lrn_rpart$train(tsk_attr, train.idx)
lrn_rpart

```

Now do the same for the random forest learner:
```{r train_ranger}
lrn_ranger
lrn_ranger$train(tsk_attr, train.idx)
lrn_ranger

```

Now that we have our models trained we can make predictions on the test data and check their performance.

### Predict
```{r predict_test}
pred_dummy  <- lrn_dummy$predict(tsk_attr, test.idx)
pred_dummy
pred_rpart  <- lrn_rpart$predict(tsk_attr, test.idx)
pred_rpart
pred_ranger <- lrn_ranger$predict(tsk_attr, test.idx)
pred_ranger

```

### Measure Performance
You can review the available measures, which are stored in the R6 dictionary `mlr_measures`.
```{r measure_list}
as.data.table(mlr_measures)[task_type == "classif"][,.(key)]
```

For simplicity we're going to only look at the accuracy score and the confusion matrix for each model.

```{r measure_models}
meas_acc <- msr("classif.acc")

pred_dummy$score(meas_acc)
pred_dummy$confusion
pred_rpart$score(meas_acc)
pred_rpart$confusion
pred_ranger$score(meas_acc)
pred_ranger$confusion
```

### SessionInfo
```{r sessionInfo}
sessionInfo()

```
