---
title: "mlr3 Intermediate"
date: "9/29/2019"
output: 
    html_document:
      toc: true
      toc_float: true
      toc_depth: 3
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup_part_one, echo = FALSE}
# recreate tasks and learners from part 1
library(ggplot2)
library(ggthemes)
library(data.table)

library(mlr3)
library(mlr3learners)
library(mlr3viz)

attrDT <- as.data.table(rsample::attrition)
tsk_attr <- TaskClassif$new(id="attrition", backend = attrDT, 
                            target = "Attrition", positive = "Yes")

lrn_dummy  <- lrn("classif.featureless")
lrn_rpart  <- lrn("classif.rpart")
lrn_ranger <- lrn("classif.ranger")
```

## Resampling

Review mlr3 resampling methods available.
```{r resampling_overview}
as.data.table(mlr_resamplings)
```

Additional resampling methods for special use cases are (or will be) available in extension package:

- [mlrspatiotemporal](https://github.com/mlr-org/mlr3spatiotemporal)

The train/predict/measure process used in the basics overview is equivalent to the "holdout" resampling method.

```{r holdout}
res_holdout <- rsmp("holdout")
res_holdout

```

The 2/3 default ratio can be changed during during object construction:
```{r holdout_2}
res_holdout <- rsmp("holdout", ratio = 0.75)
res_holdout
```

Note:

- `Instantiated: FALSE` means the resampling strategy has not been applied to a dataset yet.