---
title: "mlr3 Resampling & Benchmark"
date: "9/29/2019 (updated: `r Sys.Date()`)"
output: 
    html_document:
      toc: true
      toc_float: true
      toc_depth: 3
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
<style>
div.litegreen { background-color:#CEF6EC; border-radius: 5px; padding: 20px;}
</style>

```{r setup_part_one, echo = FALSE}
# recreate tasks and learners from part 1
library(ggplot2)
library(ggthemes)
library(data.table)

library(mlr3)
library(mlr3learners)
library(mlr3viz)

attrDT <- as.data.table(rsample::attrition)
tsk_attr <- TaskClassif$new(id="attrition", backend = attrDT, 
                            target = "Attrition", positive = "Yes")

lrn_dummy  <- lrn("classif.featureless")
lrn_rpart  <- lrn("classif.rpart", predict_type = "prob")
lrn_ranger <- lrn("classif.ranger", num.trees=15L, predict_type = "prob")
lrn_log    <- lrn("classif.log_reg", predict_type = "prob")

msr_acc <- msr("classif.acc")
msr_auc <- msr("classif.auc")
```

## Resampling

Review mlr3 resampling methods available.
```{r resampling_overview}
as.data.table(mlr_resamplings)
```

Additional resampling methods for special use cases are (or will be) available in extension package:

- [mlrspatiotemporal](https://github.com/mlr-org/mlr3spatiotemporal)

### Holdout setup

The train/predict/measure process used in the basics overview is equivalent to the "holdout" resampling method.

```{r holdout}
res_holdout <- rsmp("holdout")
res_holdout

```

The 2/3 default ratio can be changed during during object construction:
```{r holdout_2}
rs_holdout <- rsmp("holdout", ratio = 0.75)
rs_holdout
```

Note:

- `Instantiated: FALSE` means the resampling strategy has not been applied to a dataset yet.

### Holdout instantiation
There are two methods to apply the holdout strategy on a dataset:

1. manually by calling the `.$instantiate()` method on a `Task`

`res_holdout$instantiate(tsk_attr)`

2. automatically by passing the resampling object to `resample()`

`res_rpart <- resample(tsk_attr, lrn_rpart, res_holdout)`

```{r holdout_rpart}
set.seed(3334)
rs_rpart <- resample(tsk_attr, lrn_rpart, rs_holdout)
rs_rpart
rs_rpart$score(msr_auc)
```

```{r holdout_ranger}
set.seed(3334)
rs_ranger <- resample(tsk_attr, lrn_ranger, rs_holdout)
rs_ranger
rs_ranger$score(msr_auc)
```

```{r holdout_log_reg}
set.seed(3334)
rs_log <- resample(tsk_attr, lrn_log, rs_holdout)

rs_log$score(msr_auc)
```

### Holdout evaluation
```{r holdout_log_auc}
rs_log$prediction()$confusion

library(precrec)
evaluated_log = evalmod(
    scores = rs_log$prediction()$prob[, tsk_attr$positive],
    label = rs_log$prediction()$truth,
    posclass = tsk_attr$positive
)

ggplot2::autoplot(evaluated_log, curvetype="ROC")
```


## Benchmark

Benchmarking is used to compare the performance of different learners 1 or more tasks and/or using 1 or more resampling strategies.

### Design
The first step in benchmarking is to prepare a design, which is a matrix of settings to execute. Settings are made up `Task`, `Learner`, and `Resampling` objects.

For this example we're still working with a single task, `tsk_attr`.

```{r design}
rcv3 <- rsmp("repeated_cv", folds = 3, repeats = 5)

learners <- c("classif.featureless", "classif.rpart", "classif.log_reg", "classif.ranger")
learners <- lapply(learners, lrn, predict_type="prob",
                   predict_sets = c("train", "test"))

resamplings <- c(rs_holdout, rcv3)
learners
resamplings

```


```{r design_2}
design <- benchmark_grid(tsk_attr, learners, resamplings)
design

```

### Execute Benchmarks

```{r benchmark_execute, include=FALSE}
attr_bm <- benchmark(design, store_models = TRUE)
```

```{r}
attr_bm

```


### Benchmark Evaluate

```{r benchmark_evalutate}
measures <- list(
    msr("classif.auc", id = "auc_train", predict_sets = "train"),
    msr("classif.auc", id = "auc_test", predict_sets = "test")
)

tab = attr_bm$aggregate(measures)
ranks = tab[, .(learner_id, resampling_id, rank_train = rank(-auc_train), rank_test = rank(-auc_test)), by = task_id]


ranks[, .(mrank_train = mean(rank_train), mrank_test = mean(rank_test)), 
      by = .(learner_id, resampling_id)][order(mrank_test)]
```



### SessionInfo
```{r sessionInfo}
sessionInfo()

```