---
title: "mlr3 Tuning"
date: "9/29/2019 (updated: `r Sys.Date()`)"
output: 
    html_document:
      toc: true
      toc_float: true
      toc_depth: 4
      theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r setup_part_one, echo = FALSE}
# recreate tasks and learners from part 1
library(ggplot2)
library(ggthemes)
library(data.table)

library(mlr3)
library(mlr3learners)
library(mlr3tuning)
library(mlr3pipelines)
library(mlr3viz)

attrDT <- as.data.table(rsample::attrition)
tsk_attr <- TaskClassif$new(id="attrition", backend = attrDT, 
                            target = "Attrition", positive = "Yes")

lrn_dummy  <- lrn("classif.featureless")
lrn_rpart  <- lrn("classif.rpart", predict_type = "prob")
lrn_ranger <- lrn("classif.ranger", predict_type = "prob")
lrn_log    <- lrn("classif.log_reg", predict_type = "prob")

msr_acc <- msr("classif.acc")
msr_auc <- msr("classif.auc")
```

```{r validate_data, include=FALSE}
attr_newdata <- fread("../00_Data/attrition_newdata.csv")
```

# Model Optimization

## Hyperparameter Tuning

To maximize model performance a data-driven optimization of hyperparameters (= tuning) should be conducted. In order to tune a machine learning algorithm, we need to specify:

- the search space
- the optimization algorithm (aka tuning method)
- an evaluation method, i.e., a resampling strategy 
- a performance measure

Hyperparameter tuning requires the extension package <strong>`mlr3tuning`</strong>

The main components of `mlr3tuning` are the R6 classes:

- `TuningInstance`: describes the tuning problem and stores results.
- `Tuner`: base class for implementing tuning algorithms.

### `TuningInstance`
We can look at the available parameter for a learner with `$param_set`

```{r ranger_param_set}
as.data.table(lrn_ranger$param_set)[,.(id, class, lower, upper, default)]
```
We're going to only use a couple of parameters to keep it simple and quick:

- `num.trees`
- `mtry`

```{r ranger_params}
library(paradox)
tn_rng_ps <- ParamSet$new(list(
    ParamInt$new("num.trees", lower = 100, upper = 1000),
    ParamInt$new("mtry", lower = 5, upper = 15)
))

tn_rng_ps
```

Next we define how to evaluate the performance by defining:

- `resampling strategy`
- `performance measure`

```{r tune_perf}
cv3 <- rsmp("cv", folds = 3)
msr_auc <- msr("classif.auc")
```

The last step is to determine the 'budget' to solve this tuning instance by defining a `Terminator`.
Available `Terminators` are:

- `TerminatorClockTime`: terminate after a given time
- `TerminatorEvals`: terminate after a given amount of iterations
- `TerminatorPerfReached`:  terminate after a specific performance is reached
- `TerminatorCombo`: A combination of the above in an ALL or ANYTHING fashion

For this example we're going to give a budget of 20 iterations.

```{r }
# term_secs = term("clock_time")
# term_secs$param_set
# term_secs$param_set$values$secs = 5

evals20 =  term("evals", n_evals = 20)

rng_tn_instance <- TuningInstance$new(
    task = tsk_attr,
    learner = lrn_ranger,
    resampling = cv3,
    measures = msr_auc,
    param_set = tn_rng_ps,
    terminator = evals20
)

rng_tn_instance
```

### `Tuner` class

The following algorithms are currently implemented in mlr3tuning:

- Grid Search (`TunerGridSearch`)
- Random Search (`TunerRandomSearch`) (Bergstra and Bengio 2012)
- Generalized Simulated Annealing (`TunerGenSA`)

```{r tune_class}
tuner_rng <- tnr("grid_search", resolution = 10)
```

Start tuning... (output suppressed)

```{r tuning, include=FALSE}
future::plan("multiprocess")
set.seed(0820)
rng_tune_res <- tuner_rng$tune(rng_tn_instance)

```

### Review Results
```{r}
rng_tn_instance$archive(unnest = "params")[,c("num.trees", "mtry", "classif.auc")]
```

You could now use the optimal values to train a learner and make predictions on new data.

`lrn_ranger$param_set$values = rng_tn_instance$result$params`

`lrn_ranger$train(tsk_attr)`

We're not going to make predictions on new data at this point, but here's what the learner looks like before and after updating the parameters:
```{r tuned_learner}
lrn_ranger

lrn_ranger$param_set$values = rng_tn_instance$result$params

lrn_ranger

```

And after training...
```{r train_ranger_tuned}
lrn_ranger$train(tsk_attr)

lrn_ranger
```

### Make Predictions
```{r}
pred_ranger = lrn_ranger$predict_newdata(tsk_attr, attr_newdata)

pred_ranger$score(c(msr_acc, msr_auc))

```

## Autotuner


`future::plan("multiprocess")`

```{r}
future::plan("multiprocess")
tuner = tnr("random_search")
meas_auc <- msr("classif.auc")


at_ranger = AutoTuner$new(
    learner = lrn("classif.ranger", predict_type = "prob"),
    resampling = cv3,
    measures = meas_auc,
    tune_ps = tn_rng_ps,
    terminator = evals20,
    tuner = tuner
)

```


```{r}
at_ranger

```



```{r include=FALSE}
set.seed(3334)
at_ranger$train(tsk_attr)


```

```{r}

at_ranger

at_pred = at_ranger$predict_newdata(tsk_attr, attr_newdata)

at_pred$score(c(msr_acc, msr_auc))

```

### SessionInfo
```{r sessionInfo}
sessionInfo()

```
