---
title: "synthpop @ St. Louis R User Group"
output: html_document
---

**R library synthpop.**
https://cran.r-project.org/web/packages/synthpop/vignettes/synthpop.pdf

Kay Apperson, PhD.

October, 2017

```{r echo=FALSE}
# suppressPackageStartupMessages()
list.of.packages <- c("synthpop")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]
if (length(new.packages)) install.packages(new.packages)

library(synthpop)
```

**Examples of synthpop.**

SD2011: The Social Diagnosis project (Council for Social Monitoring 2011) 

which aims to investigate objective and subjective quality of life in Poland.

```{r}
head(SD2011)
```
Using a subset of the variables.
```{r}
vars <- c("sex", "age", "edu", "marital", "income", "ls", "wkabint")
ods <- SD2011[, vars]
head(ods)
```

The synthesis is just one line of code `syn(ods, seed = my.seed)`.

sds is the resulting **sythesized** data set.
```{r}
my.seed <- 17914709
sds.default <- syn(ods, seed = my.seed)
sds.default
```

**sds.default$syn** is the dataframe we're looking for.
```{r}
head(sds.default$syn)
str(sds.default$syn)

names(sds.default)

sdsdefault <- sds.default$syn
```

Note: We could view data in csv from RTVS or write CSV in R using `write.csv(MyData, file = "MyData.csv")`

**Force parametric. **
```{r}
sds.parametric <- syn(ods, method = "parametric", seed = my.seed)

syn(ods,)
sds.parametric$method
```
Comparing a variable
```{r}
compare(sds.default, ods, vars = "income")
compare(sds.parametric, ods, vars = "income")
```

***
**Extended examples.**

*Ex1.* Change the sequence of variables, and also skip a few of them.
```{r}
sds.selection <- syn(ods, visit.sequence = c(1, 2, 6, 4, 3), seed = my.seed, drop.not.used = TRUE)
sds.selection
```

*Ex2.* Build parameter.matrix.corrected without synthesizing first.

Then use parameter.matrix.correct in another synthesis.

1. Build.
```{r}
visit.sequence.ini <- c(1, 2, 5, 6, 4, 3)
method.ini <- c("sample", "ctree", "ctree", "polyreg", "", "ctree", "")
sds.ini <- syn(data = ods, visit.sequence = visit.sequence.ini, method = method.ini, m = 0, drop.not.used = FALSE)

sds.ini$predictor.matrix

predictor.matrix.corrected <- sds.ini$predictor.matrix
predictor.matrix.corrected["marital", "ls"] <- 0
predictor.matrix.corrected
```

2. Use.
```{r}
sds.corrected <- syn(data = ods, visit.sequence = visit.sequence.ini,
                 method = method.ini, predictor.matrix = predictor.matrix.corrected,
                 seed = my.seed)

```

*Ex 3.* Define numeric missing data encodings, e.g. -9999, etc. In this case, it's -8.

In this case, income can be NA or -8 when missing.
```{r}
sds.income <- syn(ods, cont.na = list(income = c(NA, -8)), smoothing = list(income = "density"), seed = NA)
sds.income
```

*Ex 4.* Restricted values according to business logic.

Males must be 18 or older to be legally married. 

Build rules.marital and use it in the parameter "rules" in syn().

1. Build.
```{r}
M18.ods <- table(subset(ods, age < 18 & sex == "MALE", marital))
M18.default <- table(subset(sds.default$syn, age < 18 & sex == "MALE", marital))
M18.parametric <- table(subset(sds.parametric$syn, age < 18 & sex == "MALE", marital))
cbind("Observed data" = M18.ods, CART = M18.default, Parametric = M18.parametric)

rules.marital <- list(marital = "age < 18 & sex == 'MALE'")
rvalues.marital <- list(marital = "SINGLE")
```

2. Use two ways, default and parametric.
```{r echo=FALSE}
sds.rmarital <- syn(ods, rules = rules.marital, rvalues = rvalues.marital, seed = my.seed)
sds.rmarital.param <- syn(ods, rules = rules.marital, rvalues = rvalues.marital, method = "parametric", seed = my.seed)
```

3. A summary table can be produced as follows.
```{r}
rM18.default <- table(subset(sds.rmarital$syn, age < 18 & sex == "MALE", marital))
rM18.parametric <- table(subset(sds.rmarital.param$syn, age < 18 & sex == "MALE", marital))
cbind("Observed data" = M18.ods, CART = rM18.default, Parametric = rM18.parametric)
```

***
**Synthetic data analysis**
```{r}
ods$wkabint <- as.character(ods$wkabint)
ods$wkabint[ods$wkabint == "YES, TO EU COUNTRY" | ods$wkabint == "YES, TO NON-EU COUNTRY"] <- "YES"
ods$wkabint <- factor(ods$wkabint)
ods$income[ods$income == -8] <- NA
```

Synthesize 5 datasets/syntheses using m = 5.
```{r}
sds <- syn(ods, method = "ctree", m = 5, seed = my.seed)
summary(ods)
summary(sds)
```

Use msel to show one or more syntheses.
```{r}
summary(sds, msel = 2)
summary(sds, msel = 1:5)

compare(sds, ods, vars = "income")
compare(sds, ods, vars = "ls", msel = 1:3)
```

**Build models on original data and syntheses.**

There are glm.synds() and lm.synds() functions to use for syntheses.
1. The glm model for ods.
```{r}
model.ods <- glm(wkabint ~ sex + age + edu + log(income), family = "binomial", data = ods)
model.ods
summary(model.ods)
```

2. The glm for sds.
```{r}
model.sds <- glm.synds(wkabint ~ sex + age + edu + log(income),
             family = "binomial", data = sds)
model.sds
summary(model.sds)

# next to each other.
summary(model.ods)
summary(model.sds)
```
